\documentclass[11pt]{article}

\input{macros}

\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc,shapes.geometric,
                decorations.pathreplacing}
\usepackage[numbers,sort&compress]{natbib}

\sloppy
\emergencystretch=3em

\title{Vegas: A Domain-Specific Language for\\
       On-Chain Strategic Protocols}
\author{Elazar Gershuni\\\texttt{elazarg@gmail.com}}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
A smart contract that implements an auction, a voting procedure, or
any other multi-party economic mechanism is two things at once: a
program that runs on a public, adversarially ordered execution layer,
and a strategic game played by rational participants.  The contract
is usually written, audited, and deployed as the former; reasoning
about its incentive properties is left to a parallel pen-and-paper
exercise on the latter.  Vegas is a domain-specific language and
compiler that treats both views as outputs of a single specification.

A Vegas program describes a finite protocol in the natural sequential
style of a referee: who joins, who plays what, who reveals, and who
gets paid.  The compiler analyses the data flow of the specification
to build a finite \emph{event graph} --- a directed acyclic graph of
guarded writes --- and uses it as a single intermediate representation
that each backend reads in turn: (i) a Solidity (and Vyper) contract
whose execution scheduler is the dependency relation itself, with
built-in commit-reveal protection against transaction-ordering
adversaries and a timeout-driven bail-out for non-responsive
participants; (ii) a Gambit extensive-form game suitable for
equilibrium analysis; (iii) several auxiliary artifacts --- an
SMT-LIB encoding for reachability queries, a Coq/Lean witness-record
encoding of the protocol's per-step constraint structure, a
session-typed (Scribble) protocol description, and a sketch
Bitcoin/Lightning move transcript for the two-party fragment.  The expressive power of the
language is restricted by design: only finite types, no loops, no
inter-contract calls.  These restrictions are what make the whole
array of artifacts feasible to produce automatically from one source.

The report describes the language, the event-graph intermediate
representation, the automatic commit-reveal rewriting pass, the
backends, and the relationship to a companion Lean~4 library called
VegasCore.  We are explicit about the parts that are stable, the
parts that are scoped down on purpose, and the parts that we expect
to relax as the language evolves.
\end{abstract}

\tableofcontents

\input{sections/01-introduction}
\input{sections/02-language}
\input{sections/04-event-graph}
\input{sections/05-commit-reveal}
\input{sections/06-solidity}
\input{sections/07-gambit}
\input{sections/08-other-backends}
\input{sections/09-liveness}
\input{sections/10-vegascore}
\input{sections/11-evaluation}
\input{sections/12-related-work}
\input{sections/13-discussion}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
