\section{The Solidity and Vyper Backends}
\label{sec:solidity}

The EVM backends translate the EventGraph into a deployable smart
contract.  Both languages target the same compilation pass via a
shared EVM intermediate representation (\TT{evm.EVMIR}); only the
final pretty-printer differs.  This section describes the actual
generated code rather than an idealised version: the bail-out
mechanism in particular is simpler than the language-level story
might suggest, and a few of its consequences belong in the
open-questions list rather than the guarantees list.

\subsection{Threat model}
\label{sec:solidity:threats}

The compiled contract is meant to survive a transaction-ordering
adversary --- a builder or competing searcher who can choose which
pending transactions to include and in what order --- and to remain
liveness-preserving when a participant stops responding.  It is not
designed against:

\begin{itemize}\itemsep1pt
\item \emph{Censorship.}  An adversary who indefinitely excludes a
  participant's transactions is modeled as forcing a timeout, which
  marks the participant as \TT{bailed} and triggers the language-level
  quit handler.
\item \emph{Cross-contract interactions.}  The contract is a closed
  system; no calls into other contracts are emitted (the only
  outbound action is a direct ETH transfer to a registered
  participant address during withdrawal).
\item \emph{Gas-cost games.}  Gas is a real-world cost that affects
  which moves are rational, but the analytical backends ignore it.
  Bounded per-action gas envelopes (e.g.\ via
  GASTAP~\citep{Albert2020Gastap}) would be needed to bring gas into
  the strategic model; this is planned but not implemented.
\item \emph{Cryptographic primitives.}  The contract relies on
  keccak-256 being binding and hiding.  Domain separation and
  per-deployment binding are in the hash composition
  (\S\ref{sec:solidity:commit-reveal}), but the strength of the
  primitive itself is taken as given.
\end{itemize}

The remainder of this section describes the construction.  The
design's looser parts --- especially the single global timeout
window --- are flagged inline as they come up.

\subsection{Storage layout}

The contract's storage is derived directly from the EventGraph.
There are five families of slots.

\paragraph{Per-node completion.}  A two-level mapping
\TT{actionDone[role][actionId]} records, for each role and each
node, whether the node has fired.  A symmetric mapping
\TT{actionTimestamp[role][actionId]} stores the block timestamp at
which it fired.  Each node is assigned a deterministic
topological index; one
\TT{uint256 constant ACTION\_\textit{Role}\_\textit{n}} is
emitted per node, where \textit{n} is the node's stable identifier
(the phase index for nodes lifted directly from source, a fresh
counter for nodes introduced by the commit-reveal expansion ---
see \S\ref{sec:eventgraph}) and the constant's \emph{value} is
its topological index.  Per-action functions in the contract are
named \TT{move\_\textit{Role}\_\textit{topo}}, using the
topological index for uniqueness across the whole contract.

\paragraph{Role registry.}  A mapping
\TT{roles[address] => Role} associates EOAs with the roles they
have joined, and a per-role \TT{address\_\textit{Role}} slot
stores the registered address.  A small \TT{enum Role} carries one
constructor per role plus a \TT{None} sentinel.

\paragraph{Game state.}  For each \emph{written field} in the
EventGraph the contract stores the field's value.  A field with
\PUBLIC visibility uses a slot of the field's native type
(\TT{int256}, \TT{bool}, \ldots).  A field with \COMMIT
visibility uses two slots: a \TT{bytes32} commitment hash filled
in at commit time, and a separate native-type slot for the
cleartext value filled in at reveal time.  Each slot is
shadowed by a \TT{done\_\textit{slot}: bool} flag, so the
contract can distinguish ``never written'' from ``written to
the default value''.

\paragraph{Per-role bookkeeping.}  Three booleans per role:
\TT{done\_\textit{Role}} (the role has joined),
\TT{claimed\_\textit{Role}} (the role has claimed its terminal
payout), and a private \TT{bailed[\textit{Role}]} (the role has been
marked inactive after a timeout).

\paragraph{Game-wide constants.}  A constant \TT{TIMEOUT} sets the
inactivity window, in seconds (the default is 24 hours).  A
constant \TT{COMMIT\_TAG} (a hashed domain-separation tag) and the
contract's deployed address are folded into every commitment hash.

\subsection{Modifiers and the depends pattern}
\label{sec:solidity:depends}

The contract has no global step counter.  Each per-node function
is guarded by a \TT{depends} modifier per predecessor.  The
modifier does two things: it triggers a bail of the named role
if the named predecessor is still unfinished after the timeout
window has elapsed, and it conditionally requires the predecessor
to have completed.

\begin{lstlisting}[style=sol]
modifier depends(Role role, uint256 actionId) {
    if (!actionDone[role][actionId]
        && block.timestamp > lastTs + TIMEOUT) {
        bailed[role] = true;
        lastTs = block.timestamp;
    }
    if (!bailed[role]) {
        require(actionDone[role][actionId], "dependency not satisfied");
    }
    _;
}
\end{lstlisting}

The bail check is gated on \TT{!actionDone[role][actionId]}: a
role is bailed only if the specific predecessor named here is
still unfinished and the global \TT{lastTs} window has expired.
A role that has already completed its predecessor cannot be
bailed by a later caller naming a different predecessor of
theirs.  If the role is not bailed, the predecessor's completion
flag is required; if it is bailed, the \TT{require} is skipped
and the dependent action proceeds.  Two additional modifiers
complete the gating: \TT{action(role, id)} asserts that the
action has not yet fired, marks it done, runs the body, and
updates \TT{actionTimestamp} and \TT{lastTs}; \TT{by(role)}
asserts that the calling EOA holds the given role and that the
role is not itself bailed.  \TT{by} does not consult the
timeout: a fair-playing caller cannot be bailed by their own
call.

A typical per-node function looks like (taken from the generated
Monty Hall contract, where node 3 is Guest's public yield of the
chosen door):

\begin{lstlisting}[style=sol]
function move_Guest_3(int256 _d)
    public
    by(Role.Guest)
    action(Role.Guest, 3)
    depends(Role.Host, 2)
{
    require(((_d >= 0) && (_d <= 2)), "domain");
    Guest_d = _d;
    done_Guest_d = true;
}
\end{lstlisting}

In this Monty Hall example the function-name suffix
(topological index) and the modifier arguments (stable node
identifiers) happen to coincide --- the source is already in
topological order, and no commit-reveal expansion runs.  In
games where expansion introduces fresh ids, the two diverge:
the bidder commit \TT{move\_B1\_4(\ldots) action(Role.B1, 2)}
in the generated Vickrey contract is a typical example.
The function's \TT{depends} clauses list exactly the node's
direct predecessors in the EventGraph (after the commit-reveal
expansion); for the rest of the chain, the \TT{require} chains
through transitive predecessors via their own modifiers.
Concurrent nodes share their predecessor prefix and are otherwise
free to be called in any order.

\subsection{Commits and reveals}
\label{sec:solidity:commit-reveal}

Every COMMIT-tagged write becomes a function that takes a
\TT{bytes32} commitment and stores it; every REVEAL-tagged write
takes the cleartext value plus a salt, recomputes the commitment,
and asserts equality with the stored hash before performing the
public write.  The commitment hash is composed as

\begin{equation*}
  \mathit{keccak256}\bigl(
    \mathrm{abi.encode}(
      \textsc{COMMIT\_TAG},\
      \mathrm{address(this)},\
      \mathrm{role},\
      \mathrm{actor},\
      \mathit{keccak256}(\mathrm{payload})
    )
  \bigr).
\end{equation*}

The pieces serve distinct roles.  The \texttt{COMMIT\_TAG} provides
cross-protocol domain separation.  \TT{address(this)} ties the
commitment to this particular deployment, preventing replay across
copies of the same protocol.  The \TT{role} identifier defends
against role-swapping replays within a single deployment; the
\TT{actor} address binds the commitment to the EOA that posted it.
The payload is pre-hashed before being folded in, so the outer
hash works over a fixed-size input regardless of the parameter
list.

Vyper emits the analogous construction using the language's
\TT{assert} idiom and slightly different storage syntax.  The
generated bytecode is comparable.

\subsection{The single-window timeout and its consequences}
\label{sec:solidity:bail}

The timeout mechanism is simpler than the language-level story
might suggest, and it is worth describing exactly.

The contract maintains a single global \TT{lastTs} slot, updated
to \TT{block.timestamp} on every successful action.  Bail is
triggered only inside the \TT{depends(role, id)} modifier
(\S\ref{sec:solidity:depends}): when
\TT{actionDone[role][id]} is still false and
\TT{block.timestamp > lastTs + TIMEOUT}, that role is marked
\TT{bailed} and the window resets.  There is therefore one
sliding inactivity window shared across all participants, but the
bail is attributed precisely to roles whose specific named action
is unfinished, not to the caller and not to bystanders.

Two properties of the emitted code are worth flagging:

\begin{itemize}\itemsep1pt
\item \emph{Precision of attribution.}  The backend emits
  \TT{depends(role, id)} clauses whose role is always the owner
  of the named predecessor, so the only role a bail can mark is
  one with an outstanding action.  A role that has completed
  everything required of it cannot be bailed by a later caller,
  and a fair-playing caller cannot be bailed by their own
  attempt to act.  These are properties of the \emph{emitter} ---
  they follow from how Vegas builds the modifier list per node
  --- and are not separately checked at runtime.
\item \emph{One-way bail.}  Once \TT{bailed[\textit{R}] = true},
  the \TT{by(R)} modifier rejects any subsequent action from
  \TT{R}.  This matches the language-level treatment of
  \TT{Quit} as a definitive move.  The remaining risk is the
  ordinary one: a fair-playing role whose connection is unstable
  for longer than \TT{TIMEOUT} loses their seat even if they
  intended to play.
\end{itemize}

When a dependent action runs against a bailed predecessor, the
\TT{depends} modifier skips its \TT{require}, but it does \emph{not}
fabricate writes for the missing fields.  Their value slots and
their \TT{done\_} flags remain in their default state (zero, and
\TT{false} respectively).  Downstream code that needs to
distinguish ``never written'' from ``written to the default value''
relies on the language's \TT{opt T} typing: a field declared with
an \TT{or null} handler is typed \TT{opt T}, and the \TT{withdraw}
clause is required to inspect its \TT{done\_} flag.

The fair-play guarantee one would like --- ``every role can drive
the contract from any reachable configuration to a terminal
allocation'' --- is plausible for protocols whose \TT{withdraw}
clause is total on the public state (including \TT{done\_} flags),
but it is not proved.  Stating and proving it under the explicit
execution model is the central refinement-theorem target of the
ongoing VegasCore work; today, the test suite checks that the
\TT{withdraw} clause type-checks against the optional fields it
must handle, and that hand-played example traces reach a terminal
allocation, but no machine-checked proof is shipped.

\subsection{Withdraw and payout}

The terminal \TT{withdraw} clause is compiled into one entry
point per role.  Each role calls its own
\TT{withdraw\_\textit{Role}()} to receive their share, guarded by
a \TT{claimed\_\textit{Role}} flag to prevent double-payment.  The
payout expression is lowered to EVM IR by direct structural
translation; the body computes the expression against the public
state (including \TT{done\_} flag inspection for optional fields)
and transfers the resulting amount to the role's stored address
using a low-level \TT{call} with the computed value.

Conservation --- that all role payouts sum to the pool --- is
a design invariant of the language (\S\ref{sec:language}), but
the current compiler does not statically verify it across
arbitrary payoff expressions.  A static check (symbolic or, for
the finite-type case, by exhaustive enumeration) is on the TODO
list; until it lands, the property rests on per-program
inspection of the \TT{withdraw} clause.

\subsection{Things the backend deliberately leaves alone}

Several real concerns are not addressed by the Solidity emission
because they are scoped out of the language.  Gas costs as a
strategic factor; mempool-level censorship beyond what the bail
mechanism handles; cross-contract interactions and oracle reads;
admin keys, upgrade proxies, and pausable patterns.  Each is a deliberate restriction, in the same spirit as the
language's bounded-types-only stance.  Drawing the boundary
tightly is what keeps the analytical backends' outputs faithful
to the deployed code, and what makes a single source for two
views a defensible claim.
