\section{Connection to VegasCore}
\label{sec:vegascore}

VegasCore is a companion Lean~4 library
\citep{deMouraUllrich2021,mathlib2020} that formalizes the
event-graph calculus underlying Vegas's source language and
proves the structural properties Vegas's IR relies on.  This
section says what VegasCore is, what it mechanizes, and how the
two artifacts fit together.  Vegas and VegasCore are independent
codebases: the compiler is useful without the library, and the
library is useful without the compiler.  What VegasCore adds is
a mechanized account of the EventGraph-side facts that make
source-level game analysis a well-defined target for backend
refinement.

\subsection{What VegasCore is}

VegasCore formalizes the core
\TT{commit}/\TT{reveal}/\TT{sample}/\TT{let}/\TT{ret} calculus
underlying Vegas's source language.  Its central object is an
\TT{EventGraph} type elaborated from a checked program
(\TT{WFProgram}); five well-formedness conditions are
statically encoded as fields of \TT{WFProgram}:
\TT{WFCtx} (scoped visibility of the typing context),
\TT{FreshBindings} (each variable is bound at exactly one
program point), \TT{RevealComplete} (every committed value is
revealed on every non-quit path), \TT{NormalizedDists}
(every \TT{sample} carries a probability distribution that
sums to one), and \TT{Legal} (every commit site has at least
one legal value in every reachable configuration).  From a checked
program the library derives a probabilistic \TT{Machine}, native
strategy carriers, kernel-game presentations, and bridges to
factored-observation stochastic games (FOSGs)~\citep{KovarikEtAl2022FOSG}
and extensive-form games (EFGs); on top of these it states and
proves a small inventory of structural and strategic theorems.

\subsection{What VegasCore mechanizes}
\label{sec:vegascore:theorems}

The theorem inventory is the centre of the library; it is also
the answer to the question ``where do the structural properties
Vegas's IR claims live?''

\begin{itemize}\itemsep2pt
\item \emph{Definitional lowering.}  Surface
  \TT{VegasLang} programs lower definitionally to the typed
  core \TT{VegasCore}; the nullable-yield compiler is provably
  equal to the canonical lowering pointwise.  Concrete programs
  are checked by reflection.  This is what lets a
  language-level program be treated as a core-calculus object
  without an intermediate trust step.
\item \emph{Event-graph adequacy.}  The elaboration
  \TT{VegasCore $\to$ EventGraph} produces a finite ranked DAG
  whose terminal-configuration outcome agrees with the program's
  declared payoff expressions.  This grounds the claim that the
  EventGraph Vegas's frontend builds and the kernel-game
  presentations VegasCore proves theorems about denote the same
  protocol.
\item \emph{Frontier diamond and preservation.}  Distinct
  enabled events in an event graph preserve each other's
  enablement and commute extensionally.  This is the formal
  counterpart of Vegas's claim
  (\S\ref{sec:eventgraph:frontier}) that concurrent nodes can be
  scheduled in either order without changing the resulting
  configuration.
\item \emph{Commit/reveal information barrier.}  No reveal node
  is on the frontier of any configuration whose graph still has
  an unfinished prior commit.  This is the property Vegas's
  commit-reveal expansion pass (\S\ref{sec:cr:barrier}) is
  designed to produce; in VegasCore it is proved once for the
  canonical event graph and inherited by every graph in
  commit-reveal normal form.
\item \emph{Trace-to-public agreement.}  The blocked-trace pure
  and PMF-behavioral kernel games derived from the event graph
  agree, on outcome kernels and expected utilities, with their
  terminal-public-state projections.  This is the technical
  underpinning of the claim that Vegas's Gambit output and a
  hypothetical trace-game analysis would agree.
\item \emph{Finite Kuhn realization.}  The bounded event-graph
  FOSG satisfies the mixed-to-behavioral strategy equivalence;
  this transports through native/FOSG equivalence to Vegas's
  kernel-game presentation.  In particular, a mixed analysis
  done over per-history pure strategies and a behavioural
  analysis done over per-information-set strategies yield the
  same outcome distributions.
\item \emph{EFG presentation.}  The canonical FOSG-to-EFG bridge
  preserves payoff-vector outcome laws.  This is the formal
  basis for treating the Gambit \TT{.efg} output as a faithful
  presentation of the program's strategic content.
\item \emph{Stochastic-refinement transport.}  A stochastic step
  refinement plus a compatible block-law lift yields an
  expected-utility-preserving morphism from a backend kernel
  game into the canonical event-graph kernel game; Nash
  equilibria pull back along this morphism.  This is the
  scaffolding on which a future backend-refinement proof (the
  Solidity-to-event-graph direction in particular) would be
  built.
\end{itemize}

The first two items are what Vegas's IR and commit-reveal
expansion pass deliver structurally; they are stated above so a
Vegas reader can see that the structural properties claimed in
\S\ref{sec:eventgraph}--\S\ref{sec:commit-reveal} are not just
``true by inspection of the rewrite'' but mechanically proved
about the canonical form the rewrite produces.  The remaining
items mechanize what one would otherwise reason about by hand
when treating a Vegas program as a game.

\subsection{Shared vocabulary}

The two projects use the same names for the things they have in
common:

\begin{center}
\begin{tabular}{ll}
\toprule
Vegas (this compiler) & VegasCore (Lean) \\
\midrule
EventGraph                  & EventGraph \\
node / nodeId               & event / node id \\
guard (\TT{scope}, \TT{expr}) & per-node guard $R$ \\
\COMMIT / \REVEAL / \PUBLIC writes & hidden + commit/reveal node kinds \\
frontier (\TT{FrontierMachine}) & frontier \\
\TT{random}                 & \TT{sample} \\
\TT{withdraw}               & \TT{Ret} \\
\bottomrule
\end{tabular}
\end{center}

A reader of either document who has read this table is equipped
to cross-reference the other.

\subsection{Where they diverge, and why}

Each artifact has features the other does not.

\paragraph{Surface keywords and macros.}  VegasCore's surface
language exposes the five core constructors directly; Vegas's
surface adds the user-facing
\TT{yield}/\TT{join}/\TT{commit}/\TT{reveal}/\TT{random}
keywords, macros, the \TT{or split/burn/null} handler forms, and
the \TT{||} per-query handler.  These all lower into the core
constructors at IR construction time; they exist in Vegas
because they make specifications readable, not because they
enlarge the semantic universe.

\paragraph{The commit-reveal expansion pass.}  Vegas's IR runs
an explicit rewrite (\S\ref{sec:commit-reveal}) that introduces
commit-reveal pairs for concurrent public writes.  VegasCore
models programs whose graphs are \emph{already} in commit-reveal
normal form: its calculus has commit and reveal as primitive
constructors and never needs a rewrite.  VegasCore's barrier
theorem (\S\ref{sec:vegascore:theorems}) holds for any program
in normal form.  The open piece is a matching \emph{equivalence}
theorem: that the Vegas rewrite preserves the strategic content
of the pre-rewrite graph.  The expected route is a bisimulation
between the pre- and post-rewrite frontier machines under the
actor's view; we have not stated it formally.

\paragraph{The operational handler layer.}  Quit handlers,
bail-out timeouts, and the structural choices in the Solidity
backend (\S\ref{sec:solidity:bail}) are first-class in Vegas;
VegasCore presently treats \TT{Quit} as a strategic move only,
without modelling the deadline mechanism.  A
backend-refinement chapter in VegasCore covers a generic notion
of operational refinement that could carry the bail mechanism,
but the Vegas-to-VegasCore mapping of the timeout subgame is
not yet written down.

\paragraph{Static obligations.}  VegasCore's \TT{WFProgram}
carries the five well-formedness obligations as fields;
Vegas's typechecker and IR constructor enforce the same
conditions but spread across the typechecker, the EventGraph
constructor, and the commit-reveal pass without exposing them
under a single uniform name.  Lifting them under VegasCore-style
names is a planned refactor.

\paragraph{Backends.}  Vegas emits Solidity, Vyper, Gambit,
SMT-LIB, Scribble, MAID, GraphViz, the Diamond DAG witness
records (\S\ref{sec:other:diamond}), and a sketch
Bitcoin/Lightning transcript; VegasCore emits no backends ---
it is a library of definitions and theorems.

\subsection{What this means for using Vegas today}
\label{sec:vegascore:status}

Vegas's design premise is that game-theoretic reasoning about a
program transfers, modulo the stated execution model, to the
deployed contract.  The compiler structures the source so that
this transfer is well-defined: a Vegas program is a game with
public history, information sets given by the visibility
discipline, and a payoff function over the terminal state.  Two
distinct routes are then available for drawing on the transfer.

\begin{itemize}\itemsep2pt
\item \emph{Mechanical analysis on the generated artifacts.}  The
  Gambit \TT{.efg} output drives Gambit's equilibrium solvers
  and dominance checks; the SMT-LIB output drives Z3 for
  reachability and coalition queries; the Diamond DAG
  (\S\ref{sec:other:diamond}) supports per-program proofs in
  Coq or Lean.  These artifacts are produced from the same
  EventGraph that drives the deployed Solidity contract.
\item \emph{Reasoning at the source level.}  The Vegas program
  is short enough to read as a game directly --- and for the
  cases where the bundled backends are not the right tool, the
  language is structured to make hand analysis tractable.  The
  conservation property, the visibility discipline, the explicit
  handler choices, and the commit-reveal-canonical IR all bound
  what a strategic reasoner has to consider.
\end{itemize}

VegasCore is a third route, available when the user is willing
to work in Lean.  The theorem inventory above mechanically
certifies the structural facts that ground the previous two
routes, and an in-Lean equilibrium argument can be transported
through the Kuhn-realization and EFG-presentation theorems to a
Vegas program's kernel-game presentation.

\subsection{What is and is not proved today}

The current state of the connection:

\begin{itemize}\itemsep1pt
\item For Vegas programs in commit-reveal normal form, the
  semantics VegasCore mechanizes is consistent with the
  EventGraph the compiler produces: the lowering uses the same
  dependency inference and visibility discipline, and the IR-level
  invariants Vegas's constructor enforces correspond to
  VegasCore's well-formedness obligations and the structural
  theorems above.
\item The backends agree with Vegas's IR structurally: every
  backend reads the same EventGraph and lowers it
  systematically.  This is a code-review property, not a formal
  one --- there is no compiler proof.
\item There is no formal proof, today, that any specific
  backend's output refines the EventGraph in any externally
  meaningful sense.  The Solidity-to-EventGraph refinement
  proof, in particular, would close the most strategically
  important gap; it is the central new theorem proposed in the
  fellowship-funded continuation of this work.
\item The Diamond DAG (Coq/Lean) backend
  (\S\ref{sec:other:diamond}) is a different route: rather than
  prove correctness of the compiler once, it ships a typed
  certificate \emph{with each compiled program}.  Both routes
  --- verified compiler and verified compiled artifact --- are
  on the project's TODO list as separate workstreams.
\end{itemize}

In short: Vegas's transferability premise is defended today by
three things --- the design of the IR, the mechanically proved
theorems VegasCore contributes about that structure, and the
disciplined, code-reviewed structure of the backends.  A
formally certified end-to-end pipeline is the natural next
milestone, not a prerequisite for present use.
