\section{Examples and Testing}
\label{sec:evaluation}

This section describes the example library, the golden-master
test harness, and what we have learned from running them.  We
make no quantitative claim: the goal is to show that the
pipeline covers a range of protocol shapes, not to benchmark
against any baseline.

\subsection{The example library}

The repository contains thirty-nine Vegas specifications under
\file{examples/}.  They are written by hand and intended to
exercise the language across three axes: textbook game-theory
examples, smart-contract patterns, and adversarial constructions.

\begin{description}\itemsep2pt
\item[Classical games.]
  \file{Prisoners.vg} (Prisoners' Dilemma),
  \file{OddsEvens.vg} and \file{OddsEvensShort.vg} (matching
    pennies, full and minimal),
  \file{MontyHall.vg} and \file{MontyHallChance.vg} (with a
    public coin),
  \file{Centipede.vg},
  \file{UltimatumGame.vg},
  \file{RPSLS.vg} (rock-paper-scissors-lizard-Spock),
  \file{Dominance.vg},
  \file{TicTacToe.vg},
  plus tiny smoke-test programs \file{Simple.vg} and
    \file{Trivial1.vg}.

\item[Mechanism design.]
  \file{VickreyAuction.vg} (sealed-bid second-price),
  \file{EntryFeeAuction.vg},
  \file{GasPriceAuction.vg},
  \file{HiddenReserve.vg} (reserve-price auction with hidden
    reserve),
  \file{Lottery.vg}, \file{ThreeWayLottery.vg},
  \file{ThreeWayLotteryShort.vg},
  \file{ThreeWayLotteryBuggy.vg} (an intentionally-flawed
    variant used to demonstrate the equilibrium analyses
    catching a vulnerability),
  \file{CommitteeVoting.vg},
  \file{StakeAndVote.vg}.

\item[Smart-contract patterns.]
  \file{EscrowContract.vg} (buyer/seller/arbiter escrow with
    dispute),
  \file{MicropaymentChannel.vg} (two-party payment channel),
  \file{AtomicSwap.vg},
  \file{BribedArbitrator.vg} (collusion-resistance test),
  \file{ClaimFraud.vg},
  \file{DataAvailability.vg},
  \file{Insurance.vg},
  \file{StakingSlashing.vg},
  \file{PrivacyNoise.vg}.

\item[Adversarial / structural.]
  \file{Bet.vg} (simple betting),
  \file{DoubleFlipLights.vg},
  \file{CongestionRouting.vg},
  \file{RandomLeader.vg},
  \file{RumorForwarding.vg},
  \file{Puzzle.vg},
  \file{SpinTheDial.vg},
  \file{TwoRobotCorridor.vg}.
\end{description}

Most examples are between ten and forty lines of Vegas source.
\file{TicTacToe.vg} is a stress test for bounded enumeration of
board states; \file{Insurance.vg} is the longest in source
length.  The Vickrey auction is the canonical mechanism example
and the cleanest demonstration of the commit-reveal expansion
pass.

\subsection{The golden-master harness}

The test suite is structured around \emph{golden masters}:
expected backend outputs checked into the repository under
\file{src/test/resources/golden-masters/}, with one subdirectory
per backend.  For each backend $B$ and each example $E$ for
which $B$ is applicable, the test harness compiles $E$ with $B$
and asserts byte-for-byte equality with the golden master
\file{golden-masters/$B$/$E$.\textit{ext}}.  When a regression
runs, the actual output is written to \file{test-diffs/$B$/$E$.\textit{ext}}
and a diff is generated; a small Python helper
\file{copy-test-outputs.py} promotes accepted changes back to
the golden masters.

This is the same kind of testing one would do for a compiler
that emits any large text artifact: most changes to the
generated code are wrong (regressions); the few that are right
are accepted explicitly.  The harness is the project's primary
defense against unintended drift across backends.

Backends are not always applicable to all examples.  Three
common reasons:

\begin{itemize}\itemsep1pt
\item Gambit and other game-theoretic backends reject any program
  whose move alphabet is unbounded (\TT{int}); these examples
  are restricted to the operational backends in the test list.
\item Bitcoin/Lightning supports only the two-strategic-role
  subset; multi-bidder auctions are excluded.
\item MAID rejects any guard referencing a field written by
  another node (\S\ref{sec:other-backends}); guarded examples
  carry an explicit backend-disable annotation.
\end{itemize}

Each example has a small declaration listing the backends to
disable; the golden-master harness uses this list to skip an
example for a backend it cannot accept.  Byte-equal golden
masters are pinned for a representative subset of programs
covering every backend exercised by the harness (MontyHall,
OddsEvensShort, Prisoners, TicTacToe, VickreyAuction; the latter
two carry a backend-disable annotation for game-theoretic and
Lightning targets respectively).  The full example library is
exercised by the language-level passes --- parse, typecheck,
conservation enumeration, and IR construction --- on every
program.

\subsection{What we have learned}

A few patterns that the example library has surfaced.

\paragraph{Distribution semantics changes the games.}  Many of
the classical game-theory examples cannot be stated literally in
Vegas: with a fixed pool, both-defect cannot be made worse than
both-cooperate in absolute terms.  The \file{Prisoners.vg}
variant therefore shifts the outcome matrix while preserving the
Nash structure.  This is not a weakness; it follows from the
conservation property the language enforces.  It does matter,
however, for translating textbook problems faithfully.

\paragraph{Guards that look local are often contextual.}  The
Monty Hall \TT{Host.goat != Guest.d} clause is the canonical
example: it reads as a local constraint on Host's move, but it
references another role's prior write.  This pattern is easy to miss when writing Vegas by hand.  The MAID
backend's restriction is what brought this pattern into focus
(the MAID encoding ignores it silently).

\paragraph{Liveness handlers are non-trivial design.}  Choosing
between \TT{or split}, \TT{or burn}, and \TT{or null} for any
given action is a genuine design decision that affects both the
contract and the game.  The example library carries
explanatory comments around these choices, especially in
\file{OddsEvens.vg}, where the comment block discusses the
distinction between losing (a strategic outcome) and quitting
(a non-strategic disruption) and the payoffs that make them
distinguishable.

\paragraph{Front-running protection is the main payoff for
auctions.}  The Vickrey example, written naively as three
simultaneous public \TT{yield}s, is automatically rewritten to
a commit-reveal protocol.  Reading the Solidity output makes the
rewriting visible: each bidder gets a pair of functions
\TT{move\_$B_i$\_$c$(bytes32 \_hidden\_b)} for the commit and
\TT{move\_$B_i$\_$r$(int256 \_b, uint256 \_salt)} for the reveal
(with topological indices $c, r$), and the reveal's
\TT{depends} chain lists every other bidder's commit as a
prerequisite.  A hand-written equivalent would need careful
attention to make this dependency right; here the IR computes
it once.
