\section{Related Work}
\label{sec:related}

Vegas sits at the intersection of three lines of work --- domain
specific languages for smart contracts, automated game-theoretic
analysis, and event-structure semantics for concurrent processes.
None of these alone covers Vegas's design, and the value of the
language design is in their combination.

\subsection{Smart-contract domain-specific languages}

Several existing DSLs target contract correctness.
Marlowe~\citep{Lamela2020Marlowe} offers a small financial
contract calculus for Cardano with formal semantics in
Isabelle/HOL; its scope is multi-party financial agreements and
its proofs are about safety and conservation, not about strategic
properties of the players.
Scilla~\citep{Sergey2019Scilla} provides a typed intermediate
language for Zilliqa with Coq-verified semantics and a strict
separation of communication and computation; like Marlowe, its
guarantees concern functional behavior of the contract rather
than the game it implements.
Move~\citep{blackshear2019move} introduces linear resource types
to make ``how much can leave the contract'' a property of the
type system.

Vegas differs in two structural ways.  First, the source
language describes a \emph{game played by the contract},
not just a contract, and the compiler emits a game-theoretic
artifact (a Gambit \TT{.efg}) alongside the executable code.
Second, the language is deliberately narrower than these
alternatives: no recursion, no token transfers, no arbitrary
external calls.  This narrowness is what gives the multi-backend
story its consistency.

Verification tools that check smart contracts against temporal
or safety specifications --- VerX~\citep{Permenev2020Verx},
KEVM~\citep{park2018formal}, Certora~\citep{CertoraWhitepaper2022}
--- are complementary.  Their question is ``does this bytecode
satisfy this specification''; Vegas's question is ``which
specification?''  Strategically-aware specifications are exactly
what Vegas's Gambit/MAID/SMT artifacts are.

\subsection{Game-theoretic analysis of smart contracts}

The MEV literature~\citep{Daian2020Flash,qin2022quantifying}
documents extensive strategic activity on existing chains:
front-running, sandwich attacks, transaction reordering for
profit.  These analyses are case studies of deployed contracts,
performed after the fact and largely by hand.  Mitigation work
has typically taken two routes: implementation-level cryptographic
defenses for commit-reveal patterns~\citep{breidenbach2018enter}
and runtime alternatives such as private transaction
pools~\citep{Weintraub2022FlashbotsAuction}.  Vegas's
contribution is to build a specific subclass --- observation-before
action on concurrent public moves --- into the language: the
mitigation is a structural property of the generated contract,
not a discipline applied by humans.

\subsection{Probabilistic programming languages}
\label{sec:related:ppl}

Probabilistic programming languages such as
Church~\citep{Goodman2008Church},
Anglican~\citep{wood2014anglican},
WebPPL~\citep{Goodman2014WebPPL},
Pyro~\citep{Bingham2019Pyro}, and
Gen~\citep{cusumano2019gen}
make distributions first-class and let the programmer write
generative models that the runtime then runs inference over.  The
languages share two primitives that Vegas's randomness story
inherits.  First, \texttt{sample}: drawing a value from a
named distribution, with the program semantics being a measure
over execution traces.  Second, \texttt{observe} (or
\texttt{assume}): conditioning the trace on a logical
proposition holding at the sample site.  Together these give
the equation \(\TT{sample}\ x\sim D;\ \TT{observe}\ \varphi(x)
\equiv\ x\sim (D \mid \varphi)\), which is the standard
PPL view of Bayesian conditioning.

Vegas adopts this view at the surface and in the IR.  A binding
of the form \TT{(x: T $\sim$ D) where $\varphi(x)$} denotes
\(x \sim (D \mid \varphi)\): the analysis backends project
self-only guards into a renormalised prior (MAID) or condition
per-trace by filtering surviving values during enumeration
(Gambit).  The two operational realisations of the same
conditioning math correspond to the static (MAID CPD) and
dynamic (Gambit trace) styles of inference in the PPL
literature.  At the IR layer, distributions are recorded
per-node (\texttt{SampleSpec}: a finite-support PMF together
with an entropy source), so a backend that wants to compute
posterior probabilities reads the same metadata used to draw
moves.

There is a stylistic parallel with Vegas's commit-then-guard
pattern.  A commit fixes a value privately; a later guard or
reveal constrains it.  When the value is committed with a
distribution annotation, the pair behaves like
\(\TT{sample}\ x; \TT{assume}\ \varphi(x)\) under PPL
semantics: the value is drawn from $D$ and constrained at the
reveal site.  The mechanism is cryptographic (a hash binds the
committed value) where a PPL would be denotational (the
posterior is computed by the inference engine), but the
declarative shape is the same.

What Vegas does not take from the PPL line of work is at least
as instructive.  Vegas has no continuous distributions: every
random variable lives in a bounded finite type, so distributions
are list-of-pairs PMFs with rational weights.  It has no
inference engine: posterior computation is done by the analysis
backends (Gambit, MAID) on finite trees, not by a general-purpose
sampler over arbitrary programs.  It has no soft conditioning
(\texttt{factor}, weighted observe): guards are hard predicates
and an unsatisfiable guard is a static error, not a low-weight
trace.  It has no recursive samplers or stochastic loops: the
event graph is finite by language design.  These restrictions
let the analysis backends terminate by exhaustive enumeration
rather than approximating with sampling; the PPL community
generally relaxes them and pays for it with Monte Carlo
inference.

\subsection{Logic programming, games, and probability}
\label{sec:related:lp}

A second declarative tradition sits adjacent to the PPL one and
is worth comparing against on its own terms.  It splits cleanly
into two strands: \emph{game-aware} logic programming, in which
the role-vs-Nature distinction is a primitive of the language,
and \emph{probabilistic} logic programming, in which all
nondeterminism is Nature.

The game-aware strand has two main representatives.  Poole's
Independent Choice Logic~\citep{Poole1997ICL} partitions atoms
into \emph{choice alternatives} owned by agents, with Nature's
alternatives carrying probabilities and strategic agents'
alternatives left open for an equilibrium operator to fill in.
The mapping into Vegas is partial.  The VegasCore calculus has
five primitive constructors (\TT{ret}, \TT{letExpr}, \TT{sample},
\TT{commit}, \TT{reveal}); two of them are recognisable in ICL
and the third is the divergence point.  \TT{sample x $\sim$ D}
corresponds directly to an ICL Nature choice over the support of
$D$.  \TT{commit x@who where R} corresponds to a role-owned
choice alternative, with the guard $R$ restricting the legal
alternatives --- modulo a feature ICL does not have, that
\TT{commit}'s output is visible only to \TT{who}.  \TT{reveal}
has no ICL analogue at all: ICL has no primitive that promotes a
hidden atom to public after the fact, and the visibility lattice
on Vegas variables (\TT{pub} versus \TT{hidden\,who}) is
external to ICL's possible-worlds semantics.

Every Vegas program denotes a finite extensive-form game with
information sets, and every such EFG is encodable in ICL by
indexing each role's choice set on the prefix of atoms
observable to that role and adding an information-set
measurability constraint on strategies.  The encoding is
non-local: each \TT{commit} or \TT{reveal} alters the index list
of every downstream choice set.  The visibility lattice on
VegasCore contexts is a static discipline enforced at the type
level; the corresponding ICL property is a meta-condition on
strategies, not a property of the program text.  The
general-game-playing language
GDL-II~\citep{ThielscherGDLII}, extending the original
GDL~\citep{LovePerkinsEtAl2008GDL}, makes the same distinction
operationally: a distinguished \TT{random} role whose moves are
stipulated uniform over its \TT{legal} set, ordinary roles whose
moves are strategic, and a \TT{sees} predicate for imperfect
information.  GDL-II describes games turn by turn through
\TT{legal(R,M)} and \TT{does(R,M)} predicates over a state
fluent; Vegas's EventGraph is the asynchronous analogue, with
dependency rather than a global step counter ordering moves, and
with commit-reveal as a first-class node kind rather than a
derived information-flow pattern.  GDL-III~\citep{ThielscherGDLIII}
extends GDL-II with epistemic predicates \TT{knows} and
\TT{common\_knowledge} that may appear inside \TT{legal} and
\TT{terminal}.  This is an orthogonal direction to Vegas's:
GDL-III lets game \emph{rules read} epistemic state inferred from
the observation trace, while Vegas lets game \emph{bindings
create} epistemic state via the \TT{commit}/\TT{reveal} pair and
the visibility lattice on contexts.  Either language can encode
the other's idiom by re-routing through its native primitive ---
a Vegas \TT{reveal} corresponds to emitting \TT{sees} for every
role in GDL-III, and a GDL-III knowledge-conditioned \TT{legal}
clause corresponds, in Vegas, to a guard over a previously
revealed witness --- but the primitives chosen are different.  ICL is closer to Vegas in
flavour --- both are declarative, both make the agent-owning-a-
choice-set construct primary --- while GDL-II is closer in
intent, as a description language for whole games.

The probabilistic strand --- Sato's distribution
semantics~\citep{Sato1995DistributionSemantics},
ProbLog~\citep{DeRaedt2007ProbLog},
P-log~\citep{BaralGelfondRushton2009PLog} --- annotates logical
facts with probabilities and defines a measure over the
program's stable models or proof trees.  These languages are
not game description languages: they have no agent primitive,
and a \TT{random} fact owned by ``the Matcher'' is no
different from a \TT{random} fact owned by Nature.
P-log in particular is excellent at the Bayesian-update
question --- the cleanest way to state ``given that Host opened
door $1$ after Guest picked $0$, what is the posterior on the
car location'' is a one-line P-log query --- and Vegas's PPL
story (\S\ref{sec:related:ppl}) is the natural place to relate
that machinery.  But for the strategic question, these
languages can describe one strategy profile at a time and
require a meta-layer over strategies to talk about equilibrium.
DTProbLog~\citep{VandenBroeck2010DTProbLog} adds a single-agent
decision-theoretic layer to ProbLog and is honest about the
restriction; extending it to multi-agent equilibrium is, in
spirit, what Vegas does on the IR side, taking the ICL/GDL-II
role distinction as the right primitive on which to build.  Concurrent constraint
programming~\citep{Saraswat1991CCP} supplies the right semantic
shape for that asynchrony: an action fires when its
\emph{ask} (dependencies derivable) succeeds and then
\emph{tells} its output, which is structurally the firing rule
for an EventGraph node.

Two further connections are worth flagging even though Vegas
does not pursue them.  First, the pool's conservation law has
the shape of a linearity constraint and would sit naturally in a
linear logical framework~\citep{CervesatoPfenning2002CLF,
LopezPfenningPolakow2005LolliMon}: deposits as linear resources,
\TT{withdraw} as their exact consumption.  Vegas instead checks
conservation as an integrity property of the generated artifact.
Second, equilibrium computation is genuinely outside the
fixed-point theory of standard LP: a Nash operator is not
monotone in the lattice that gives stable models, so a logic
programming presentation of Vegas would either delegate
equilibrium to an external solver (as ASP-with-optimisation
does) or commit to a richer game-theoretic semantics on top.

The framing this comparison suggests --- which we record here as
a stylistic claim, not a formal correspondence --- is that
Vegas's source semantics is approximately \emph{concurrent
constraint logic programming with role-indexed choice
alternatives (Nature being one role whose alternatives carry
probabilities), a visibility lattice on bindings exposed through
\TT{commit}/\TT{reveal}, linear resources for the pool, and
default rules on a non-derivability of a deadline atom}.  The
\TT{commit}/\TT{reveal} pair is a \emph{primitive} of the source
calculus rather than a derived information-flow pattern: what
the EVM backend does to it (hash, store, check on reveal) is a
lowering of the primitive into a setting that does not natively
support hidden bindings, and what the Gambit backend does
(translate to information sets) is a different lowering of the
same primitive.  The unusual part of Vegas is therefore the
target, not the source semantics.

Table~\ref{tab:lp-features} summarises the comparison along
the axes that matter for Vegas's design.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lcccccccc}
\toprule
                                & ICL & GDL-II & GDL-III & ProbLog & DTProbLog & P-log & CCP & Vegas \\
\midrule
Role vs.\ Nature is primitive   & \cmk & \cmk & \cmk & \xmk & \tmk\textsuperscript{a} & \xmk & \tmk\textsuperscript{b} & \cmk \\
Observation primitive (\TT{sees}) & \tmk & \cmk & \cmk & \xmk & \xmk & \xmk & \xmk & \tmk \\
Epistemic predicates in guards  & \xmk & \xmk & \cmk & \xmk & \xmk & \xmk & \xmk & \xmk \\
Visibility lattice on bindings  & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \cmk \\
\TT{commit}/\TT{reveal} as primitive & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \cmk \\
Bayesian conditioning (\TT{observe}) & \tmk & \xmk & \xmk & \cmk & \cmk & \cmk & \xmk & \cmk \\
Asynchronous DAG firing         & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \cmk & \cmk \\
Linear resource / conservation  & \xmk & \xmk & \xmk & \xmk & \xmk & \xmk & \tmk & \cmk \\
Equilibrium operator in language & \tmk\textsuperscript{c} & \tmk\textsuperscript{c} & \tmk\textsuperscript{c} & \xmk & \xmk & \xmk & \xmk & \cmk \\
\bottomrule
\end{tabular}
\caption{Feature comparison.  \cmk\ = native primitive;
\tmk\ = encodable but not primitive; \xmk\ = absent.
\textsuperscript{a}DTProbLog has a single decision agent, not
a role partition.
\textsuperscript{b}CCP has agents only in the sense of
concurrent processes, with no Nature distinction.
\textsuperscript{c}GDL/GGP and ICL leave equilibrium to an
external solver layer rather than fixing a semantics; Vegas
commits to particular equilibrium notions through its MAID and
Gambit backends.}
\label{tab:lp-features}
\end{table}

\subsection{Multi-agent influence diagrams and game representations}

MAID~\citep{koller2003multi} provides a factored representation
of multi-agent decisions as a Bayesian network with decision and
utility nodes.  We discussed the MAID backend in
\S\ref{sec:other-backends}; the relevant limit is that plain
MAIDs cannot encode context-dependent action sets, which makes
them an awkward target for Vegas games whose guards are
non-trivial.  Constrained-MAID extensions exist in the literature
but are not standardized.

The Gambit project~\citep{McKelveyMcLennanTurocy2006Gambit} and
OpenSpiel~\citep{LanctotEtAl2019OpenSpiel} provide tools for
analyzing extensive-form games once one is given.  Vegas adds
the path from a programmer's specification to such a game.

\subsection{Event structures and dependency graphs}

The EventGraph is a particular flavour of event
structure~\citep{NielsenPlotkinWinskel1981,Winskel1986EventStructures}
specialized to the protocol setting --- typed, finite,
visibility-aware, with explicit commit/reveal node kinds.  The
program-dependence graph~\citep{ferrante1987program} and
static-single-assignment form~\citep{cytron1991efficiently} are
the immediate technical antecedents on the program-analysis side:
EventGraph edges are computed by the same dependency analysis
those works pioneered, only specialized to the
\TT{commit}/\TT{reveal} discipline.

\subsection{Compositional game theory and IF logic}

Compositional Game Theory~\citep{GhaniHedgesWinschel2018CGT} and
the work it builds on develop a categorical view of open games
that compose like programs.  IF logic~\citep{HintikkaSandu1989,
MannSanduSevenster2011IFL} provides a logical apparatus for
imperfect-information games in which informational independence
is a first-class construct.  Vegas's per-program Diamond DAG
encoding (\S\ref{sec:other:diamond}) has a sympathetic flavour
--- a player's choice schedule is a function on prior witnesses,
which is the IF-logic notion of a Skolem function for an
existential under independence --- but the connection is at the
level of stylistic resonance, not a formal correspondence we
have spelled out.  The companion VegasCore
(\S\ref{sec:vegascore}) does this work on the side of native
kernel-game semantics rather than open-game composition.

\subsection{Multiparty session types}

Multiparty session types~\citep{HondaYoshidaCarbone2008MPST}
specify the legal sequencing of messages in a multi-party
protocol; Castellan and Yoshida's recent work
\citep{CastellanYoshidaSessionGame} relates session types to
game semantics directly.  Vegas's Scribble backend uses session
types as a target, projecting the EventGraph's public-message
subset onto a global type.  The relationship is more
``information set $\subseteq$ session type than the reverse'':
session types record which messages are legal in a given
prefix, but they do not directly record information sets.  A
fuller comparison is future work.

\subsection{Refinement and operational correctness}

The bail-out subgame and the Solidity scheduling layer share
the shape of an implementation refining a more abstract
specification, in the sense of Lynch and
Vaandrager~\citep{LynchVaandrager1995}.  Vegas's companion
formalization VegasCore states a stochastic-step refinement
theorem on the abstract event-graph machine.  A formal proof
that Vegas's Solidity backend refines its EventGraph is on the
project's TODO list; this is precisely the gap the fellowship
proposal addresses.
